Response diversity is a major driver of temporal stability in complex food webs

### 2024-08-29 paper/manuscript.Rmd
Add repo link to manuscript
Fix typos and clarify formulations in M&M

---
title: Response diversity is a major driver of temporal stability in complex food webs
author: "Alain Danet (1), Sonia Kéfi (2,3), Thomas F. Johnson (1), Andrew P. Beckerman (1)"
date: ""
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: true
    spacing: double
    number_sections: false
  officedown::rdocx_document:
    number_sections: false
    base_format: "bookdown::word_document2"
always_allow_html: true
bibliography: references.bib
link-citations: true
csl: nature.csl
#csl: ecology-letters.csl
toc: false
linestretch: 2.0
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{hyperref}
  - \hypersetup{backref=true}
  - \usepackage{setspace}\doublespacing
---

@@

We further found that asynchrony is driven more by portfolio effects than by
compensatory dynamics (resp. $r_\delta$= `r ras` and `r rac`). In more details, response
diversity has strong positive effects on asynchrony, mainly through portfolio
effects ($r_\delta$= `r rsd`), but also through a weaker positive effect on compensatory
dynamics ($r_\delta$= `r rcd`). All other studied properties were found to have relatively
much weaker effects on asynchrony. Food web structural properties – namely
average trophic level, connectance and average interaction strength – have
consistent negative effects on asynchrony, through both portfolio and
compensatory effects (Fig. \@ref(fig:sem)a). Indeed, as average trophic level and connectance
increase compensatory effects are reduced (resp. $r_\delta$= `r rct` and `r rcc`, Fig.
\@ref(fig:sem)a), while average interaction strength has a small negative effect
on portfolio effects ($r_\delta$= `r rsi`). Species richness has a positive
effect on both portfolio and compensatory effects (resp. $r_\delta$=`r rsr` and
`r rcr`). Finally, as expected, environmental stochasticity has little effect on
asynchrony.

@@

Deriving total effects from the sum of direct and indirect effects in the
structural equation model (See Method), we showed that both environmental
stochasticity and species response diversity have by far the strongest total
effects on community stability (resp. $r_\delta$= `r rtsenv` and `r rtsdiv`,
Fig. \@ref(fig:sem)b). Species richness and average trophic level have similar
small but positive total effects on community stability ($r_\delta$=`r rtsr`),
but through opposite pathways. While species richness increases community
stability through a strong positive effect on asynchrony ($r_\delta$=`r rtar`),
it also weakly decreases population stability ($r_\delta$=`r rtpr`). In
contrast, average trophic level increases community stability through strong
positive effect on population stability ($r_\delta$=`r rtpt`), but is dampened
by a negative total effect on asynchrony ($r_\delta$=`r rtat`). Interestingly,
average interaction strength also has a positive effect on population stability
and a negative one on asynchrony (resp. $r_\delta$= `r rtpi` and `r rtai`). This
results in a total negative effect on community stability ($r_\delta$=
`r rtsi`). Connectance, instead, has a negative effect on community stability
($r_\delta$= `r rtsct`) through negative effects on both asynchrony and
population stability (resp. $r_\delta$=`r rtact` and `r rtpct`).

@@

We simulated the dynamics of complex food-webs using the allometric bioenergetic
model [@yodzis_body_1992; @brose_allometric_2006; @delmas_simulations_2017]. This model
is widely used to simulate the dynamics of complex ecological communities
because of the simplicity  and because it can be parametrized using  the
metabolic theory of ecology [@brown_toward_2004]. The model describes species
biomass dynamics over time with two master equations, one for the primary
producers (eq. \@ref(eq:prod)) and one for the consumers (eq.\@ref(eq:cons)):

@@

The consumers gain biomass by consuming resources (i.e. their preys) at a rate
that depends on their mass specific metabolic rate ($x_i$),  maximum consumption
rate ($y_i$) and on the functional response ($F_ij$) describing how the rate of
consumption of a consumer $i$ on a resource $j$ vary with the biomass of this resource.
Consumers then lose biomass by being consumed, $e_ij$ being the assimilation
efficiency of the consumer $j$ on the resource $i$. Consumers also lose biomass
over time through metabolic losses ($-x_iB_i$) and by natural mortality due to
environment ($-d_iB_i$). The primary producers gain biomass over time with a
growth rate ($r_i$, eq. \@ref(eq:prod)) and a functional response ($G_i$)
describing how the growth of the producers vary with their biomass. They lose
biomass by being consumed by consumers and by natural mortality.

@@

We added a stochastic natural mortality rate to the species dynamics,
representing a fluctuating environment. Based on the literature and previous
models [@vasseur_environmental_2007], we assumed that natural mortality rates
scales inversely with the species body mass ($d_i = d_0M_i^{-1/4}$), as all
other physiological parameters. We used a basal natural mortality rate of $d_0 =
.4$, as done in previous studies[@vasseur_environmental_2007].

We introduced environmental stochasticity by considering stochastic variation of
mortality rates. Doing so allows to consider stochasticity without modifying
species interactions themselves
[@vasseur_environmental_2007]. The stochastic part of the mortality rates
($\epsilon_{i, t}$) follow a normal distribution 0 centered
($\epsilon_{i, t} \sim N(0,\sigma^2_e)$) with a variance ($\sigma^2_e$). The
mortality rates of each species at the time $t$ equals to $d_{i,t} = d_i
e^{\epsilon_{i, t}}$, which ensured that the mortality rates never become
negative [@vasseur_environmental_2007]. We
simulated $\epsilon$ using an Ornstein-Uhlenbeck process, which is a modified
Brownian motion where the stochastic values tend to come back to the central
value (i.e. $\epsilon = 0$) such as $\epsilon_t$ varies according to this
stochastic differential equation: $d\epsilon_t = (0 - \epsilon_t)d_t + \sigma_e
dW_t$, where $dW_t$ is a Brownian motion. The strength of environmental
stochasticity was then controlled by $\sigma_e$.

Response diversity was controlled by the correlation among species stochastic
mortality rates ($\rho_{ij, i \neq j}$) such as response diversity is maximal
when species mortality rates are uncorrelated ($\rho = 0$) and null when species
mortality rates are perfectly correlated ($\rho = 1$) [@ripa_food_2003;
@vasseur_environmental_2007; @gouhier_synchrony_2010]. So the species mortality
rates were a product of the stochastic mortality rates generated by the
Ornstein-Uhlenbeck process and a variance-covariance matrix such as:

@@

We numerically solved the stochastic differential equation systems during 2000
timesteps and collected the last 500 timesteps, similarly to previous studies
[@brose_allometric_2006]. We set $d_t = .1$, and $d_t = .05$ when instability
was detected during the simulations due to stiff changes in biomass. We set
extinction threshold to $10^{-6}$ and if an extinction happened in the last 500
timesteps, we continued to run the simulation for another 1000 timesteps until
there was no extinction events during the last 500 timesteps. When we found
disconnected species at the end of the simulation (i.e. a primary producer
without consumer or a consumer without prey), we set their biomass to 0 and
re-run the simulation for another 1000 timesteps. Except in the case of
disconnected species, we re-ran simulations with species starting biomass equal
to their biomass at the last timestep of the previous simulation. Supplementary
analysis showed that alternative choices of simulation processes, such as longer
simulations, rebuilding food-webs (i.e. resetting consumer preferences,
recomputing species bodymass according to the new food-web) after removing
disconnected species did not affect our results (Fig. S1). The food-web
generation and simulations were done using the `Julia` package
`EcologicalNetworkDynamics.jl`[@lajaaiti_ecologicalnetworksdynamicsjl_2024].

@@

We measured food-web properties that have been linked with stability in the
literature.  We measured total biomass, species richness, connectance, average
trophic level weighted by biomass, average omnivory and average interaction
strength. Food-web properties were measured at each of the 500 timesteps and
then averaged, except connectance and trophic level that were static.
Connectance was computed as $C = L / S(S-1)$, $L$ being the number of trophic
links and $S$ being the number of species in the community. Species trophic
level was computed recursively from the bottom of the food-web, such as the
trophic level of a consumer was equal to the average trophic level of its
resources added to 1, the trophic level of primary producers being set to 1
[@christensen_ecopath_1992]. Average trophic level was then equal to:
$1/\hat{B}_i\sum_{i}\mathrm{TL}_i$, $\mathrm{TL}_i$ being the trophic level and
$\hat{B}_i$ the average biomass of the species $i$. The degree of omnivory of a
consumer was computed  such as the sum of squares of its resource trophic levels
weighted by the relative preference of the consumer for each resource: $\sum_i
(\mathrm{TL}_i - \hat{\mathrm{TL}})^2 \omega_i/ \sum \omega_i$
[@christensen_ecopath_1992]. Interaction strength was computed was quantified as
the biomass fluxes going from a resource to a consumer, such as: $I = x_i y_i
B_k F_{ij}$ (eq. \@ref(eq:cons)), and averaged over time. Our way of measuring
interaction strength is very similar to previous metrics used in previous
studies, at the difference here that we average fluxes over time and not the
maximal interaction strength [@mccann_weak_1998]. Finally, the average
interaction strength of a community was computed as the average interaction
strength at the exclusion of null interaction strengths (i.e. excluding absence
of trophic interactions).

@@

with $S_{com} = \mu_{tot} / \sigma_{tot}$, $S_{pop} = \mu_{tot} / \sum_i \sigma_i$,
$\phi = \sum_i \sigma_i / \sigma_{tot}$.

$\mu_{tot}$ and $\sigma_{tot}$ being respectively average total biomass and standard
deviation of total biomass.

To further get a mechanistic insight about the effects of stressors and food-web
structure on stability, we partitioned asynchrony into portfolio effects and
compensatory dynamics stemming from species interaction[@zhao_biodiversity_2022]
($CPE_{int}$). The measurement of portfolio effects has been a topic of
debate[@loreau_biodiversity_2021;@zhao_biodiversity_2022]. Doak's
definition[@doak_statistical_1998] defines portfolio effects as the product of
statistical averaging effects ($SAE$) and compensatory effects arising from
response diversity[@zhao_biodiversity_2022; @doak_statistical_1998]
($CPE_{env}$). It then quantities the portfolio effect ($PFE = SAE \times
CPE_{env}$) emerging from independent species fluctuations, i.e. the statistical
averaging effect ($SAE$), weighted by the effects of response diversity of
species to environmental fluctuations, such as portfolio effects are dampened if
species have correlated responses.

@@

Using a structural equation model, we disentangled the complex effects of
food-web structure, environmental stochasticity and response diversity on the
different facets of temporal stability of biomass (i.e. population stability,
asynchrony, portfolio effects and compensatory dynamic effects). We tested the
effects of food-web structure (average interaction strength, average trophic
level, connectance) on population stability and asynchrony partitions (portfolio
and compensatory effects). Grounded in theoretical ecology, we expected that
higher average trophic level, lower interaction strength, lower
connectance result in higher population stability and lower asynchrony[@mccann_weak_1998; @mccann_reevaluating_1997;
@mccann_diversitystability_2000; @fahimipour_compensation_2017]. We further
added simulation parameters as control
variables because higher Predator-Prey Mass Ratio and predator interference were
also expected to increase population stability [@brose_allometric_2006].
Finally, we tested the effects of environmental stochasticity and response
diversity on population stability and asynchrony partitions, although there were
predicted to mainly affect to population stability (i.e. by increasing amplitude
of population fluctuations) and asynchrony (i.e. by generating independent
population fluctuations) respectively. We completed the structural equation
model with the mathematical relationship linking compensatory dynamics due to
species interactions ($CPE_{int}$) and statistical averaging effects ($SAE
\times CPE_{env}$) to asynchrony, and from asynchrony and population stability
to community stability.

Prior to test the structural equation model, we logged all the stability
components to transform their relation from multiplicative to additive (eq.
\@ref(eq:scom2)). We ensured that all the linear models composing the structural
equation model presented low multicollinearity (Variance Inflation Factor < 3,
Table S4), although interaction strength and species richness were highly
correlated (Fig. S5). The structural equation model using the R package
`PiecewiseSEM`[@lefcheck_piecewisesem_2016]. We then bootstrapped (N = 100) the
SEM to obtain confidence intervals on the estimated effects and computed the sum
of the direct and indirect effects using `semEff` R
package[@murphy_semeff_2021]. In the main text, we reported standardised
coefficients which were obtained by scaling the coefficients by the standard
deviation of the response and predictor variable. Finally, we reported only the
direct standardised coefficients that had an absolute value above $0.05$,
because almost all linked were statistically significant given the high numbers
of simulations.

In a second analysis, we tested how food-web structure, response diversity, and
environmental stochasticity modulate stability-species richness relationships.
We modelled temporal stability of community biomass according to food-web
structure: species richness, weighted average trophic level, connectance and
average interaction strength, as well as environmental stochasticity and
response diversity. We further included the two-way interactions between species
richness and food-web structure, between species richness and response
diversity, between species richness and environmental stochasticity. Finally, we
added the three-way interactions between species richness, food-web structure
and response diversity, between species richness, food-web structure and
environmental stochasticity. We added predator interference, predator-prey mass
ratio as control variables. We further added the ID of the food-web as a random
intercept. We then used the linear equation of the model and associated slope
coefficients to predict the shape of the relationship between community
stability and species richness according to a gradient of response diversity and
of food-web structure. To do so, we summed all the coefficients involving
species richness in the model, i.e. the one, two, and three-way terms. The
values of the other variables were set to the values indicated in the Fig.
\@ref(fig:pred). The VIF were inferior to 3, indicating low multicollinearity
(Table S5), we checked the distribution of the residuals (Fig. S6) using
`DHARMa` R package. The model was implemented using the R package
`glmmTMB`[@brooks_glmmtmb_2017].

>>>>

---
title: Response diversity is a major driver of temporal stability in complex food webs
author: "Alain Danet (1), Sonia Kéfi (2,3), Thomas F. Johnson (1), Andrew P. Beckerman (1)"
date: ""
output:
  officedown::rdocx_document:
    number_sections: false
    base_format: "bookdown::word_document2"
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: true
    spacing: double
    number_sections: false
always_allow_html: true
bibliography: references.bib
link-citations: true
csl: nature.csl
#csl: ecology-letters.csl
toc: false
linestretch: 2.0
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{hyperref}
  - \hypersetup{backref=true}
  - \usepackage{setspace}\doublespacing
---

@@

We further found that asynchrony is driven more by portfolio effects than by
compensatory dynamics (resp. $r_\delta$= `r ras` and `r rac`). In more details, response
diversity has strong positive effects on asynchrony, mainly through portfolio
effects ($r_\delta$= `r rsd`), at the expense of compensatory
dynamics ($r_\delta$= `r rcd`). All other studied properties were found to have relatively
much weaker effects on asynchrony. Food web structural properties – namely
average trophic level, connectance and average interaction strength – have
consistent negative effects on asynchrony, through both portfolio and
compensatory effects (Fig. \@ref(fig:sem)a). Indeed, as average trophic level and connectance
increase compensatory effects are reduced (resp. $r_\delta$= `r rct` and `r rcc`, Fig.
\@ref(fig:sem)a), while average interaction strength has a small negative effect
on portfolio effects ($r_\delta$= `r rsi`). Species richness has a positive
effect on both portfolio and compensatory effects (resp. $r_\delta$= `r rsr` and
`r rcr`). Finally, as expected, environmental stochasticity has little effect on
asynchrony.

@@

Deriving total effects from the sum of direct and indirect effects in the
structural equation model (See Method), we showed that both environmental
stochasticity and species response diversity have by far the strongest total
effects on community stability (resp. $r_\delta$= `r rtsenv` and `r rtsdiv`,
Fig. \@ref(fig:sem)b). Species richness and average trophic level have similar
small but positive total effects on community stability ($r_\delta$= `r rtsr`),
but through opposite pathways. While species richness increases community
stability through a strong positive effect on asynchrony ($r_\delta$= `r rtar`),
it also weakly decreases population stability ($r_\delta$= `r rtpr`). In
contrast, average trophic level increases community stability through strong
positive effect on population stability ($r_\delta$= `r rtpt`), but is dampened
by a negative total effect on asynchrony ($r_\delta$= `r rtat`). Interestingly,
average interaction strength also has a positive effect on population stability
and a negative one on asynchrony (resp. $r_\delta$= `r rtpi` and `r rtai`). This
results in a total negative effect on community stability ($r_\delta$=
`r rtsi`). Connectance, instead, has a negative effect on community stability
($r_\delta$= `r rtsct`) through negative effects on both asynchrony and
population stability (resp. $r_\delta$= `r rtact` and `r rtpct`).

@@

We simulated the dynamics of complex food-webs using the allometric bioenergetic
model [@yodzis_body_1992; @brose_allometric_2006; @delmas_simulations_2017]. This model
is widely used to simulate the dynamics of complex ecological communities
because of the simplicity of its parametrisation, using the
metabolic theory of ecology [@brown_toward_2004]. The model describes species
biomass dynamics over time with two master equations, one for the primary
producers (eq. \@ref(eq:prod)) and one for the consumers (eq.\@ref(eq:cons)):

@@

The consumers gain biomass by consuming resources (i.e. their preys) at a rate
that depends on their mass specific metabolic rate ($x_i$),  maximum consumption
rate ($y_i$) and on the functional response ($F_{ij}$) describing how the rate
of consumption of a consumer $i$ on a resource $j$ vary with the biomass of this
resource. Consumers then lose biomass by being consumed, $e_{ij}$ being the
assimilation efficiency of the consumer $j$ on the resource $i$. Consumers also
lose biomass over time through metabolic losses ($-x_iB_i$) and natural
mortality ($-d_iB_i$). The primary producers gain biomass over time with a
growth rate ($r_i$, eq. \@ref(eq:prod)) and a functional response ($G_i$)
describing how the growth of the producers vary with their biomass. They lose
biomass by being consumed by consumers and by natural mortality.

@@

We added a stochastic natural mortality rate to the species dynamics,
representing a fluctuating environment. Based on the literature and previous
models[@vasseur_environmental_2007; @hatton_linking_2019], we assumed that natural mortality rates
scales inversely with the species body mass ($d_i = d_0M_i^{-1/4}$), as all
other physiological parameters. We used a basal natural mortality rate of $d_0 =
0.4$, as a previous study[@vasseur_environmental_2007].

We introduced environmental stochasticity by considering stochastic variation in
mortality rates. Doing so allows to consider stochasticity without modifying
species interactions themselves[@vasseur_environmental_2007]. The stochastic part of the mortality rates
($\epsilon_{i, t}$) follow a normal distribution 0 centered
($\epsilon_{i, t} \sim N(0,\sigma^2_e)$) with a variance ($\sigma^2_e$). The
mortality rates of each species at the time $t$ equals to $d_{i,t} = d_i
e^{\epsilon_{i, t}}$, which ensured that the mortality rates never become
negative[@vasseur_environmental_2007]. We
simulated $\epsilon$ using an Ornstein-Uhlenbeck process, a modified
Brownian motion where the stochastic values tend to come back to the central
value (i.e. $\epsilon = 0$) such as $\epsilon_t$ varies according to this
stochastic differential equation: $d\epsilon_t = (0 - \epsilon_t)d_t + \sigma_e
dW_t$, where $dW_t$ is a Brownian motion. The strength of environmental
stochasticity was then controlled by $\sigma_e$.

Response diversity was controlled by the correlation among species stochastic
mortality rates ($\rho_{ij, i \neq j}$) such as response diversity is maximal
when the stochastic component of species mortality rates ($\epsilon_{i, t}$)
are uncorrelated ($\rho = 0$), and response diversity is null when the
stochastic component of species mortality rates are perfectly
correlated[@ripa_food_2003; @vasseur_environmental_2007; @gouhier_synchrony_2010] ($\rho = 1$).
So the species mortality rates were a product of the stochastic mortality rates
generated by the Ornstein-Uhlenbeck process and a variance-covariance matrix
such as:

@@

We numerically solved the stochastic differential equation systems during at
least 2000 timesteps and collected the last 500 timesteps, similarly to previous
studies [@brose_allometric_2006]. We numerically solve the system with $\Delta t
= 0.1$ or $\Delta t = 0.05$ when instability was detected during the simulations
due to stiff changes in biomass. We set extinction threshold to $10^{-6}$ and if
an extinction happened in the last 500 timesteps, we continued to run the
simulation for another 1000 timesteps until
there was no extinction events during the last 500 timesteps. When we found
disconnected species at the end of the simulation (i.e. a primary producer
without consumer or a consumer without prey), we set their biomass to 0 and
re-run the simulation for another 1000 timesteps. Except in the case of
disconnected species, we re-ran simulations with species starting biomass equal
to their biomass at the last timestep of the previous simulation. Supplementary
analysis showed that alternative choices of simulation processes, such as longer
simulations, rebuilding food-webs (i.e. resetting consumer preferences,
recomputing species bodymass according to the new food-web) after removing
disconnected species did not affect our results (Fig. S1). The food-web
generation and simulations were done in Julia by developing an extension of the package
`EcologicalNetworkDynamics.jl`[@lajaaiti_ecologicalnetworksdynamicsjl_2024].

@@

We measured food-web properties that have been linked with stability in the
literature. We measured total biomass, species richness, connectance, average
trophic level weighted by biomass, and average interaction
strength. Food-web properties were measured at each of the 500 timesteps and
then averaged, except connectance and trophic level that were static.
Connectance was computed as $C = L / S(S-1)$, $L$ being the number of trophic
links and $S$ being the number of species in the community. Species trophic
level was computed recursively from the bottom of the food-web, such as the
trophic level of a consumer was equal to the average trophic level of its
resources added to 1, the trophic level of primary producers being set to 1
[@christensen_ecopath_1992]. Average trophic level was then equal to:
$1/\hat{B}_i\sum_{i}\mathrm{TL}_i$, $\mathrm{TL}_i$ being the trophic level and
$\hat{B}_i$ the average biomass of the species $i$. Interaction strength was computed was quantified as
the biomass fluxes going from a resource to a consumer, such as: $I = x_i y_i
B_k F_{ij}$ (eq. \@ref(eq:cons)), and averaged over time. Our way of measuring
interaction strength is very similar to previous metrics used in previous
studies, at the difference here that we average fluxes over time and not the
maximal interaction strength [@mccann_weak_1998]. Finally, the average
interaction strength of a community was computed as the average interaction
strength at the exclusion of null interaction strengths (i.e. excluding absence
of trophic interactions).

@@

with $S_{com} = \mu_{tot} / \sigma_{tot}$, $S_{pop} = \mu_{tot} / \sum_i \sigma_i$,
$\phi = \sum_i \sigma_i / \sigma_{tot}$; $\mu_{tot}$ and $\sigma_{tot}$ being
respectively average total biomass and standard deviation of total biomass.

To further get a mechanistic insight about the effects of stressors and food-web
structure on stability, we partitioned asynchrony into portfolio effects and
compensatory dynamics stemming from species interactions[@zhao_biodiversity_2022]
($CPE_{int}$). The measurement of portfolio effects has been a topic of
debate[@loreau_biodiversity_2021;@zhao_biodiversity_2022]. Doak's
definition[@doak_statistical_1998] defines portfolio effects as the product of
statistical averaging effects ($SAE$) and compensatory effects arising from
response diversity[@zhao_biodiversity_2022; @doak_statistical_1998]
($CPE_{env}$). It then quantities the portfolio effect ($PFE = SAE \times
CPE_{env}$) emerging from independent species fluctuations, i.e. the statistical
averaging effect ($SAE$), weighted by the effects of response diversity of
species to environmental fluctuations, such as portfolio effects are dampened if
species have correlated responses to environmental fluctuations.

@@

Using a structural equation model, we disentangled the complex effects of
food-web structure, environmental stochasticity and response diversity on the
different facets of temporal stability of biomass (i.e. population stability,
asynchrony, portfolio effects and compensatory dynamic effects). We tested the
effects of food-web structure (average interaction strength, average trophic
level, connectance) on population stability and asynchrony partitions (portfolio
and compensatory effects). Grounded in theoretical ecology, we expected that
higher average trophic level, lower interaction strength, lower
connectance result in higher population stability and lower asynchrony[@mccann_weak_1998; @mccann_reevaluating_1997;
@mccann_diversitystability_2000; @fahimipour_compensation_2017]. We further
added simulation parameters as control
variables because higher Predator-Prey Mass Ratio and predator interference were
also expected to increase population stability[@brose_allometric_2006].
Finally, we tested the effects of environmental stochasticity and response
diversity on population stability and asynchrony partitions, although there were
predicted to mainly affect to population stability (i.e. by increasing amplitude
of population fluctuations) and asynchrony (i.e. by generating independent
population fluctuations) respectively. We completed the structural equation
model with the mathematical relationship linking compensatory dynamics due to
species interactions ($CPE_{int}$) and statistical averaging effects ($SAE
\times CPE_{env}$) to asynchrony, and from asynchrony and population stability
to community stability.

Prior to test the structural equation model, we logged all the stability
components to transform their relation from multiplicative to additive (eq.
\@ref(eq:scom2)). We ensured that all the linear models composing the structural
equation model presented low multicollinearity (Variance Inflation Factor < 3,
Table S4), although interaction strength and species richness were highly
correlated (Fig. S5). The structural equation model using the R package
`PiecewiseSEM`[@lefcheck_piecewisesem_2016]. We computed the sum
of the direct and indirect effects using `semEff` R
package[@murphy_semeff_2021]. In the main text, we reported standardised
coefficients which were obtained by scaling the coefficients by the standard
deviation of the response and predictor variable. Finally, we reported only the
direct standardised coefficients with an absolute value equal or above $0.05$,
because almost all effects were statistically significant given the high numbers
of simulations.

In a second analysis, we tested how food-web structure, response diversity, and
environmental stochasticity modulate stability-species richness relationships.
We modelled temporal stability of community biomass according to food-web
structure: species richness, weighted average trophic level, connectance and
average interaction strength, as well as environmental stochasticity and
response diversity. We further included the two-way interactions between species
richness and food-web structure, between species richness and response
diversity, between species richness and environmental stochasticity. Finally, we
added the three-way interactions between species richness, food-web structure
and response diversity, between species richness, food-web structure and
environmental stochasticity. We added predator interference, predator-prey mass
ratio as control variables. We further added the ID of the food-web as a random
intercept. We then used the linear equation of the model and associated slope
coefficients to predict the shape of the relationship between community
stability and species richness according to a gradient of response diversity and
of food-web structure. To do so, we summed all the coefficients involving
species richness in the model, i.e. the one, two, and three-way terms. The
values of the other variables were set to the values indicated in the Fig.
\@ref(fig:pred). The coefficients of the models are displayed in Fig. S3. The
VIF were inferior to 3, indicating low multicollinearity (Table S5), we checked
the distribution of the residuals (Fig. S6) using `DHARMa` R package. The model
was implemented using the R package `glmmTMB`[@brooks_glmmtmb_2017].

# Code availability

The code used to run the simulations, analyse the data and write the manuscript
is available on Github (https://github.com/alaindanet/xStressorsStabBEFW) and
archived on Zenodo (https://zenodo.org/records/13454946).
